---
title: "HanaahRTidy"
author: "Author/creator: [Khadija Zbair]"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HanaahRTidy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE, 
  error = TRUE
)
```

### Introduction

The `HanaahRTidy` package provides a collection of functions designed to simplify common data manipulation tasks.

### Installation

The package can be installed via:
- `devtools::install_github("YourGithubUsername/HanaahRTidy")`
- `install.packages("HanaahRTidy")`

The `HanaahRTidy` package makes use of the following packages:
- dplyr
- readr

Suggested packages: 
- knitr
- rmarkdown

### Function and Data Explanation

#### convert_to_factor()

- **Description**: This function converts a column with textual values into a factor variable, useful for categorical data analyses.
- **Usage**: `convert_to_factor(data, column)`
- **Arguments**:
  - `data`: A data frame containing the column to be converted.
  - `column`: The name of the column to convert (as a string).
- **Returns**: A data frame with the specified column converted to a factor.

#### identify_missing_data()

- **Description**: This function flags missing values in a dataset and provides summary statistics on the percentage of missingness for each column.
- **Usage**: `identify_missing_data(data)`
- **Arguments**:
  - `data`: A data frame containing the dataset.
- **Returns**: A data frame summarizing the missing data statistics.

#### import_data()

- **Description**: This function imports a dataset from a CSV file and returns a data frame.
- **Usage**: `import_data(file_path)`
- **Arguments**:
  - `file_path`: A string specifying the path to the CSV file.
- **Returns**: A data frame containing the data from the CSV file.

#### split_datetime()

- **Description**: This function splits a datetime column into separate date and time columns.
- **Usage**: `split_datetime(datetime_column)`
- **Arguments**:
  - `datetime_column`: A column containing datetime values.
- **Returns**: A data frame with separate date and time columns.

### Function Examples

Loading package:

```{r examples}
# Load package
library(HanaahRTidy)
```

#### convert_to_factor()

```{r}
data <- data.frame(gender = c("male", "female", "female", "male"))
converted_data <- convert_to_factor(data, "gender")
print(converted_data)
```

#### identify_missing_data()

```{r}
sample_data <- data.frame(
  A = c(1, 2, NA, 4),
  B = c(NA, NA, 3, 4),
  C = c(1, 2, 3, 4)
)
missing_data_summary <- identify_missing_data(sample_data)
print(missing_data_summary)
```

#### import_data()

```{r, eval=FALSE}
# Example usage
# Note: Replace "path/to/file.csv" with the actual path to your CSV file
# data <- import_data("path/to/file.csv")
# print(data)
```

#### split_datetime()

```{r}
datetime_values <- as.POSIXct(c("2023-01-01 12:34:56", "2023-01-02 23:45:01"))
split_datetime_result <- split_datetime(datetime_values)
print(split_datetime_result)
```
